HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-06-03 paper/main.tex
bin size
space

\titlespacing*{\section}
{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
\titlespacing*{\subsection}
{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}

\renewcommand{\paragraph}{
  \@startsection{paragraph}{4}
  {\z@}{.1ex \@plus .1ex \@minus .1ex}{-1.5em}
  {\normalfont\normalsize\bfseries}
}
\makeatother

@@

Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
or adequacy/fluency scales. However these provide no insight into possible errors,
and do not scale well with sentence length.
We argue for a semantics-based evaluation, which captures what meaning components
are retained in the MT output, providing a more fine-grained analysis of
translation quality, and enables the construction and tuning of semantics-based MT. 
We present a novel human semantic evaluation measure, Human
UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
semantic representation scheme.
%Making progress over previous work, the proposed measure takes into account
HUME covers
a wider range of semantic phenomena than previous methods and does not rely on semantic annotation
of the potentially garbled MT output. 
%itself, whose interpretation is often unclear.
We experiment with four language pairs, demonstrating HUME's broad applicability,
and report good
%as well as its reliability, reflected in its 
inter-annotator agreement rates and 
 correlation with human adequacy scores.

@@

%Could reinstate
%After a brief review (\secref{sec:background}), we describe HUME in detail
%(\secref{sec:hume}).
%Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
%document HUME's inter-annotator agreement and efficiency (time of annotation). 
%We further empirically compare HUME with direct assessment of human adequacy ratings,
%and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).

@@

{\small
\begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Scene & R & Relater \\
P & Process & C & Centre \\
\hline
\end{tabular}
}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation where leaves correspond to words and nodes to units.
  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
  Scene. Edge labels mark UCCA categories, listed  in the table.}
%which are largely ignored in this work
\end{figure}

@@

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

@@

Structural units have sub-units (children in the UCCA graph), which may themselves
either be atomic or structural.
% Ondrej thinks the following can be simply removed:
%The labels assigned to such nodes reflect the extent to which
%the relation between the sub-units is preserved in $t$.
%\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
%Scene elements, such as the predicate (main relation), participants and adverbials.}
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong\footnote{
  Three labels are used with atomic units, as opposed to two labels with structural units,
as atomic units are more susceptible to slight errors.}.
%here are three major cases where the relationship between the children might go
%wrong in translation and 
We will use the example ``man bites dog'' to illustrate typical examples of why a structural node
should be labelled as ``Bad'':
incorrect ordering (``dog bites man''), 
deletion (``man bites'') and insertion (``man bites biscuit dog''). 
%First, the components of the translation may be ordered differently from the source,
%such that the relation between them changes. For instance, if
%the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
%the semantic unit corresponding to the Scene will be marked as Bad, while
%the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
%Second, A superfluous word or phrase may be inserted into
%the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
%Third, one of the sub-units may be missing from the translation.
%For instance, if ``dog bites man'' is translated into ``dog bites''.

@@

We turn to discussing the main differences between HUME and HMEANT,
a semantic measure that measures the translation quality according to its similarity
with a reference, measures in terms of the overlap between their SRL annotations.
%Like HMEANT, HUME defines  a
%measure that decomposes over semantic units, rather than strings. 
%However, there are important differences in design decisions, which we summarize in this
%section.

@@

HMEANT focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses.
We are not aware of any empirical evaluation that demonstrates that verb argument structure,
taken alone, captures the crux of the sentence semantics meanings.
Moreover, relying only on verbal argument structures is less stable across paraphrases
and translations, as a verbal argument structure
may be translated to a nominal or adjectival argument structure
(e.g., ``after graduation'' may be translated into ``after he graduated'').
This may lead to unjustified penalties low HMEANT score, as the verb
in one structure has nothing to align to in the other.
On the other hand, HUME has been shown to be reasonably stable in an English-French
corpus study \cite{sulem2015conceptual}.

Consider the following example:

@@

There are however a number of other serious consequences on focusing on verbal
predicates. We conducted an analysis of the English UCCA Wikipedia
corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not
well supported by HMEANT.
First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which
generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
according to conservative estimates that only consider non-auxiliary instances of ``be''.
Second, nominal argument structures, which are ignored by  HMEANT are in fact highly
pervasive, appearing in 48.7\% of the sentences.
Third, linkers, that express an inter-relation between
clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
sentences, but are again ignored by HMEANT.
An example where a linker is omitted in translation in our experiments is
\equo{However, this review\dots}, which was translated to
the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
This omission will not be taken into account by HMEANT at all\footnote{Argument structures and linkers are explicitly marked in UCCA. Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger. Nominal argument structures are here Scenes whose main relation is headed by a noun.}.

@@

\paragraph{One Structure or Two.}
HUME only annotates the source, while HMEANT relies on
two independently constructed structural annotations, one for the reference and one
for the translation.
%First of all, the annotation process is different. We semantically annotate the
%source, then automatically project it to the translation, and ask bilingual
%annotators to assess whether this projection shows a good or bad translation. In
%HMEANT, they semantically annotate both the reference and the translation, then
%align these manually, and assess to what extent they match.
%We see this difference as crucial where semantic analysis is concerned, as it
Not annotating the translation is appealing as it is often impossible to assign a
semantic structure to a low quality translation.
On the other hand, HUME's annotation may be artificially boosting the perceived understandability
of the sentence by allowing access to the source.

\paragraph{Alignment.}
In HMEANT, the alignment between the reference and translation structures is a key
part of the manual annotation. If the alignment cannot be created, the 
translation is heavily penalized.
Previous experiments with HMEANT
\parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
suggested that sometimes, the structures of the reference and of an accurate translation
still diverge, for instance due to a different
interpretation of a PP-attachment, or the verb having an additional modifier in
%suggested that sometimes, the structures of the reference and candidate
%translations diverge without any good reason, for instance due to a different
%interpretation of a PP-attachment, the verb can have one additional modifier in
one of the structures. It would be desirable to allow modifications to
the SRL annotations at the alignment stage,
to avoid unduly penalization due to such spurious divergences.
%annotations also at the alignment stage
%to avoid unduly penalization due to such a spurious divergence.
The same issue is noted in \parcite{lo:wu:reliability:2014}:
the IAA on SRL dropped from 90\% to 61\% 
when the two aligned structures were from two different annotators.
HUME uses automatic (word-level) alignment, which only
serves as a cue for directing the attention of the annotators.
The user is expected to mentally correct the
alignment as needed, thus circumventing this difficulty.

@@

\paragraph{Monolingual vs. Bilingual Evaluation.}
\label{src-vs-ref}
HUME diverges from both HMEANT and from shallower automatic measures
like BLEU or METEOR, in not requiring a reference.
Instead, it compares the source directly with the output translation.
This requires us to employ bilingual annotators, but has the benefit of avoiding
using a reference. Reference translations may often be phrased in many structurally different
ways, which may again lead unduly penalization of the translation in cases where
it is a paraphrase of the reference.

>>>>

%\titlespacing*{\section}
%{0pt}{1ex plus 1ex minus 1ex}{1ex plus .2ex}
%\titlespacing*{\subsection}
%              {0pt}{1ex plus 1ex minus 1ex}{1ex plus .2ex}

\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\titlespacing{\paragraph}{1pt}{*0}{3pt}

\makeatother

@@

Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However these provide no insight into possible errors,
and do not scale well with sentence length.
We argue for a semantics-based evaluation, which captures what meaning components
are retained in the MT output, providing a more fine-grained analysis of
translation quality, and enables the construction and tuning of semantics-based MT. 
We present a novel human semantic evaluation measure, Human
UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
%Making progress over previous work, the proposed measure takes into account
HUME covers
a wider range of semantic phenomena than previous methods and does not rely on semantic annotation
of the potentially garbled MT output. 
%itself, whose interpretation is often unclear.
We experiment with four language pairs, demonstrating HUME's broad applicability,
and report good
%as well as its reliability, reflected in its 
inter-annotator agreement rates and 
 correlation with human adequacy scores.

@@

%Could reinstate
After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}).
Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document HUME's inter-annotator agreement and efficiency (time of annotation). We further empirically compare HUME with direct assessment of human adequacy ratings, and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).

@@

{\small
  \begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Parallel Scene & R & Relater \\
P & Process & C & Centre \\
\hline
\end{tabular}
}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation. Leaves correspond to words and nodes to units.
  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
  Scene. Edge labels mark UCCA categories.}
%which are largely ignored in this work
\end{figure}

@@

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate
%\footnote{Instead of relying on the source,
%the evaluation could be based on the UCCA structure for the \emph{reference}
%translation.
%More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
%}.

@@

Structural units have sub-units (children in the UCCA graph),
which are themselves atomic or structural.
% Ondrej thinks the following can be simply removed:
%The labels assigned to such nodes reflect the extent to which
%the relation between the sub-units is preserved in $t$.
%\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
%Scene elements, such as the predicate (main relation), participants and adverbials.}
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong\footnote{
  Three labels are used with atomic units, as opposed to two labels with structural units,
as atomic units are more susceptible to slight errors.}.
%here are three major cases where the relationship between the children might go
%wrong in translation and 
We will use the example ``man bites dog'' to illustrate typical examples of why a structural node
should be labelled as ``Bad'':
incorrect ordering (``dog bites man''), 
deletion (``man bites'') and insertion (``man bites biscuit dog''). 
%First, the components of the translation may be ordered differently from the source,
%such that the relation between them changes. For instance, if
%the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
%the semantic unit corresponding to the Scene will be marked as Bad, while
%the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
%Second, A superfluous word or phrase may be inserted into
%the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
%Third, one of the sub-units may be missing from the translation.
%For instance, if ``dog bites man'' is translated into ``dog bites''.

@@

We discuss the main differences between HUME and HMEANT,
a human MT evaluation metric that measures the overlap between the translation a reference
in terms of their SRL annotations.

%Like HMEANT, HUME defines  a
%measure that decomposes over semantic units, rather than strings. 
%However, there are important differences in design decisions, which we summarize in this
%section.

@@

HMEANT focuses on verbal argument structures, ignoring other pervasive phenomena such as non-verbal predicates and inter-clause relations. Consider the following example:

@@

We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not well supported by HMEANT.
First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences, according to conservative estimates that only consider non-auxiliary
instances of ``be''.
Second, nominal argument structures, ignored by HMEANT, are in fact highly
pervasive, appearing in 48.7\% of the sentences.
Third, linkers that express inter-relations between
clauses (mainly discourse markers and conjunctions) appear in 56\% of the
sentences, but are again ignored by HMEANT. As noted in our experiments, linkers
are sometimes omitted in translation, but these omissions are not taken
into consideration by HMEANT\footnote{Argument structures and linkers are explicitly marked in UCCA. Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger. Nominal argument structures are here Scenes whose main relation is headed by a noun.}.

%An example where a linker is omitted in translation in our experiments is
%\equo{However, this review\dots}, which was translated to
%the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
%This omission will not be taken into account by HMEANT at all

We are not aware of any empirical argument suggesting that verb argument structures,
taken alone, capture the crux of the sentence semantics.
Moreover, relying only on verbal argument structures is less stable across paraphrases
and translations, as a verbal argument structure
may be translated to a nominal or adjectival argument structure
(e.g., ``after graduation'' may be translated into ``after he graduated'').
This may lead to an unduly low HMEANT score, as the verb
in one structure has nothing to align to in the other.
On the other hand, UCCA has been shown to be reasonably stable in an English-French
corpus study \cite{sulem2015conceptual}.

@@

\paragraph{One Structure or Two.}
HUME only annotates the source, while HMEANT relies on
two independently constructed structural annotations, one for the reference and one
for the translation.
%First of all, the annotation process is different. We semantically annotate the
%source, then automatically project it to the translation, and ask bilingual
%annotators to assess whether this projection shows a good or bad translation. In
%HMEANT, they semantically annotate both the reference and the translation, then
%align these manually, and assess to what extent they match.
%We see this difference as crucial where semantic analysis is concerned, as it
Not annotating the translation is appealing as it is often impossible to assign a
semantic structure to a low quality translation.
On the other hand, HUME may be artificially boosting the perceived understandability
of the translation by allowing access to the source.

\paragraph{Alignment.}
In HMEANT, the alignment between the reference and translation structures is a key
part of the manual annotation. If the alignment cannot be created, the 
translation is heavily penalized.
\newcite{bojar:wu:ssst:2012} and \newcite{chuchunkov-tarelkin-galinskaya:2014:SSST-8}
argue that the structures of the reference and of an accurate translation
may still diverge, for instance due to a different
interpretation of a PP-attachment, or the verb having an additional modifier in
%suggested that sometimes, the structures of the reference and candidate
%translations diverge without any good reason, for instance due to a different
%interpretation of a PP-attachment, the verb can have one additional modifier in
one of the structures. It would be desirable to allow modifications to
the SRL annotations at the alignment stage,
to avoid unduly penalizing such spurious divergences.
%annotations also at the alignment stage
%to avoid unduly penalization due to such a spurious divergence.
The same issue is noted in \parcite{lo:wu:reliability:2014}:
the IAA on SRL dropped from 90\% to 61\% 
when the two aligned structures were from two different annotators.
HUME uses automatic (word-level) alignment, which only
serves as a cue for directing the attention of the annotators.
The user is expected to mentally correct the
alignment as needed, thus circumventing this difficulty.

@@

\paragraph{Monolingual vs. Bilingual Evaluation.}
\label{src-vs-ref}
HUME diverges from HMEANT and from shallower measures
like BLEU, in not requiring a reference.
Instead, it compares the source directly with the output translation.
This requires the employment of bilingual annotators, but has the benefit of avoiding
using a reference, which is never uniquely defined, and may thus 
lead to unjustly low scores where the translation is a paraphrase of the reference.
